\pisodm: Building Service-Oriented Applications in the Presence of Non-Functional Requirements

### 2015-03-31 JournalToSubmit2015/conclusions.tex
first version for IJSEKE

We presented \pisodm, an MDA methodology for designing and developing reliable service-based applications. 
\pisodm extends a previously defined method (called SOD-M) to include Non-Functional Requirements.
These requirements are taken into account from the early stages of the software development process.
Non-functional constraints are related to business rules associated to the behavior of the application and, in the case of service-based applications, they are also concerned with constraints imposed by the services.

Our methodology includes two CIM-level models, three PIM-level models and one PSM-level model. 
We implemented the meta-models on the Eclipse platform and we validated the approach by using an industrially inspired use case.

>>>>

We presented \pisodm, an model-driven method for designing and developing reliable service-based applications.
\pisodm extends a previously defined method (called SOD-M) to include Non-Functional Requirements.
These requirements are taken into account from the early stages of the software development process.
Non-functional constraints are related to business rules associated to the behavior of the application and, in the case of service-based applications, they are also concerned with constraints imposed by the services.

Our method includes two CIM-level models, three PIM-level models and one PSM-level model.
We implemented the meta-models on the Eclipse platform and we validated the approach by using an industrially inspired use case.

### 2015-03-31 JournalToSubmit2015/intro.tex
first version for IJSEKE

The literature stresses the need for meth\-od\-ol\-o\-gies and techniques for service oriented analysis and design 
%since they are the cornerstone in the development of meaningful service based applications
~\cite{Papazoglou2007}. 
Existing approaches maintain that the convergence of model-driven soft\-ware development, service orientation,   and  busi\-ness processes improvement are key for developing accurate  software~\cite{watson}. 
Model Driven Development (MDD)  for software systems is mainly characterized by the use of models as a product~\cite{Selic03}.
These models are successively refined from abstract specifications into actual computer programs.

@@

This paper is organized as follows.
Sec\-tion \ref{sec:relworks} summarizes the general principles of existing works for addressing NFP and associating them to service compositions.
Sections~\ref{sec:motivation}  and~\ref{sec:mmrules} introduce respectively the meta-models of the \pisodm and transformation rules. 
% Section~\ref{sec:implementation} describes the \pisodm environment that provides tools for building service based applications according to the \pisodm methodology. 
Section~\ref{sec:flyingPig} describes the experiment that we conducted for validating our methodology and discusses lessons learned. 
Finally, Section~\ref{sec:conclusions} concludes the paper and gives research perspectives.

>>>>

The literature stresses the need for meth\-od\-ol\-o\-gies and techniques for service oriented analysis and design
%since they are the cornerstone in the development of meaningful service based applications
~\cite{Papazoglou2007}.
Existing approaches maintain that the convergence of model-driven soft\-ware development, service orientation,   and  busi\-ness processes improvement are key for developing accurate  software~\cite{watson}.
Model Driven Development (MDD)  for software systems is mainly characterized by the use of models as a product~\cite{Selic03}.
These models are successively refined from abstract specifications into actual computer programs. A recent literature review study concludes there are a very scarce attention to non-functional requirements in the development of service-oriented applications using model-driven methodologies~\cite{Ameller201542}.

@@

The applicability of our proposal was tested with an industrial case study concerning risk assessment for financial companies as implemented by the ORCA System\footnote{The ORCA System is a trademark of GCP Global (www.gcpglobal.com).}. In this work we show the application of our method \pisodm to develop a service based application called \FlyingPig that provides risk assessment as a service.The models presented were generated as a result of interacting with software developers at GCP Global and trying to attend to all the non-functional requirements present in the application domain.

This paper is organized as follows.
Sec\-tion \ref{sec:relworks} summarizes the general principles of existing works for addressing NFP and associating them to service compositions.
Sections~\ref{sec:motivation} and~\ref{sec:mmrules} introduce respectively the meta-models of the \pisodm and transformation rules.
% Section~\ref{sec:implementation} describes the \pisodm environment that provides tools for building service based applications according to the \pisodm methodology.
Section~\ref{sec:flyingPig} describes the \FlyingPig case study that we develop for validating our method and discusses lessons learned.
Finally, Section~\ref{sec:conclusions} concludes the paper and gives research perspectives.

### 2015-03-31 JournalToSubmit2015/main.tex
first version for IJSEKE

\documentclass[twocolumn]{svjour3}

\usepackage{geometry}
%\geometry{letterpaper}                   % ... or a4paper or a5paper or ...
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{amssymb} 
\usepackage{epstopdf}

\usepackage{amssymb}

%\usepackage{natbib}

%%% OUR MACROS %%%
\newcommand{\COMMENT}[1]{ }

\usepackage[usenames,dvipsnames]{xcolor}

\usepackage{amsmath}
%\usepackage[thmmarks,amsmath]{ntheorem}

\newcommand{\openbox}{\leavevmode
  \hbox to.77778em{%
  \hfil\vrule
  \vbox to.675em{\hrule width.6em\vfil\hrule}%
  \vrule\hfil}}

%\theoremstyle{plain}
%\theoremheaderfont{\normalfont\bfseries}
%\theorembodyfont{\normalfont}
%\theoremseparator{}
%\theoremindent0cm
%\theoremnumbering{arabic}
%\newtheorem{algo}{Algorithm}
%
%%\theoremstyle{plain}
%%%\theoremheaderfont{\normalfont\itshape}
%%\theoremheaderfont{\normalfont\bfseries}
%%\theorembodyfont{\normalfont}
%%\theoremseparator{}
%%\theoremindent0cm
%%\theoremnumbering{arabic}
%%\theoremsymbol{\ensuremath{\openbox}} 
%%\newtheorem{example}{Example}
%
%
%\theoremstyle{plain}
%\theoremheaderfont{\normalfont\bfseries}
%\theorembodyfont{\normalfont}
%\theoremseparator{.}
%\theoremindent0cm
%\theoremnumbering{arabic}
%\theoremsymbol{\ensuremath{\Box}} 
%\newtheorem{defi}{Definition}
%
%\theoremstyle{plain} 
%\theoremsymbol{\ensuremath{\Box}} 
%\theoremseparator{.} 
%\newtheorem{prop}{Property}

\def\FlyingPig{\textsl{FlyingPig}}

@@

\usepackage{listings}

\lstset{numbers=right, numbersep=5pt, numberstyle=\tiny, stepnumber=1,escapechar=\!,columns=fullflexible,
        morekeywords={procedure,let,for,do,if,then,else,add,choose,end,while,
        true,false,rise,exception,extend,resume,to,return,function}}

\newcommand{\pisodm}[0]{$\pi$SOD-M\xspace}

@@

\title{\pisodm: A New Methodology for Building Service-Oriented Applications}\thanks{\textsl{Contact author:}\\
Martin A. Musicante\\
UFRN - Cx Postal 1647, \\
Campus Universitário Lagoa Nova \\
59078-970, Natal, RN, Brazil\\[2mm]
Telephone: +55 84 3215 3814 ext 1.208\\
Mobile phone: +55 84 9653 4331\\
Fax: +55 84 3215 3781}

\author{Khalid Belhajjame \and Valeria~de~Castro \and Umberto~Souza~da~Costa \and Javier~A.~Espinosa-Oviedo \and Martin~A.~Musicante \and Pl\'acido~A.~Souza~Neto \and Genoveva~Vargas-Solar \and Jos\'e-Luis~Zechinelli-Martini}

\institute{K. Belhajjame
               \at Universit\'e de Paris - Dauphine -- Paris, France
               \and
               V. de Castro
               \at Universidad Rey Juan Carlos -- M\'{o}stoles, Spain
               \and
               U.S.Costa \and M.A. Musicante
               \at Universidade Federal do Rio Grande do Norte -- Natal, Brazil
               \and
               J.A. Espinosa-Oviedo, J-L Zechinelli-Martini
               \at Universidad de las Am\'ericas-Puebla, LAFMIA -- Cholula, Mexico
               \and 
               P. Souza Neto 
               \at Instituto Federal de Educa\c{c}\~{a}o, Ci\^{e}ncia e Tecnologia do Rio Grande do Norte -- Natal, Brazil
               \and
               G. Vargas-Solar
               \at CNRS, LIG-LAFMIA, Saint Martin d'H\`eres, France}

@@

\begin{abstract}

Specifying non-functional requirements (NFRs) is a complex task, being usually dealt with on the later phases of the software process.
The late inclusion of NFRs in the development may compromise the quality of the deployed application.
This paper presents \pisodm, a methodology and associated tools that 
\textit{(i)}  allows the early specification of non-functional requirements in a principled way: users are abstracted away from low level details; 
\textit{(ii)} embraces the MDA philosophy, generating models (code) whenever possible.
The pro\-po\-sed solution has been utilized in the context of an industrial case study.
\end{abstract}

@@

\section{Applying \pisodm: The \FlyingPig\ use case}
\label{sec:flyingPig}

>>>>

\documentclass{ws-ijseke}
\usepackage{xspace}

\newcommand{\pisodm}[0]{$\pi$SOD-M\xspace}
\def\FlyingPig{\textsl{FlyingPig}}

@@

@@

\markboth{Authors' Names}
{Instructions for Typing Manuscripts (Paper's Title)}

%%%%%%%%%%%%%%%%%%%%% Publisher's Area please ignore %%%%%%%%%%%%%%%
%
\catchline{}{}{}{}{}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{\pisodm: Building Service-Oriented Applications in the Presence of Non-Functional Requirements}

\author{Khalid Belhajjame}
\address{Universit\'e de Paris - Dauphine -- Paris, France\\
\email{kbelhajj@googlemail.com}
}

\author{Valeria~de~Castro} 
\address{Universidad Rey Juan Carlos -- M\'{o}stoles, Spain\\
\email{valeria.decastro@urjc.es}
}

\author{Umberto~Souza~da~Costa} 
\address{Universidade Federal do Rio Grande do Norte -- Natal, Brazil\\
\email{umberto@dimap.ufrn.br}
}

\author{Javier~A.~Espinosa-Oviedo} 
\address{Universidad de las Am\'ericas-Puebla, LAFMIA -- Cholula, Mexico\\
\email{javiera.espinosa@gmail.com}
}

\author{Martin~A.~Musicante} 
\address{Universidade Federal do Rio Grande do Norte -- Natal, Brazil\\
\email{mam@dimap.ufrn.br}
}

\author{Pl\'acido~A.~Souza~Neto} 
\address{Instituto Federal de Educa\c{c}\~{a}o, Ci\^{e}ncia e Tecnologia do Rio Grande do Norte -- Natal, Brazil\\
\email{placido.neto@ifrn.edu.br}
}

\author{Genoveva~Vargas-Solar} 
\address{CNRS, LIG-LAFMIA, Saint Martin d'H\`eres, France\\
\email{genoveva.vargas@imag.fr}
}

\author{Jos\'e-Luis~Zechinelli-Martini}
\address{Universidad de las Am\'ericas-Puebla, LAFMIA -- Cholula, Mexico\\
\email{joseluis.zechinelli@udlap.mx}
}

@@

\begin{history}
\received{(Day Month Year)}
\revised{(Day Month Year)}
\accepted{(Day Month Year)}
%\comby{(xxxxxxxxxx)}
\end{history}

\begin{abstract}
Specifying non-functional requirements (NFRs) is a complex task, being usually dealt with on the later phases of the software process.
The late inclusion of NFRs in the development may compromise the quality of the deployed application.
This paper presents \pisodm, a method and associated tools that
\textit{(i)}  allows the early specification of non-functional requirements in a principled way: users are abstracted away from low level details;
\textit{(ii)} embraces the MDA philosophy, generating models (code) whenever possible.
The pro\-po\-sed solution has been utilized in the context of an industrial and real case study.
\end{abstract}

@@

\section{Applying \pisodm: The \FlyingPig\ case study}
\label{sec:flyingPig}

### 2015-03-31 JournalToSubmit2015/modelingFlyingPig.tex
first version for IJSEKE

We validated our methodology by developing a use-case concerning risk assessment for financial companies as implemented by the ORCA System\footnote{The ORCA System is a trademark of GCP Global (www.gcpglobal.com).}.
Risk assessment is implemented by an interactive business process based on the exchange of a series of questionnaires intended to evaluate the risks implied by the client business practice.
Examples of business practices are: the conditions and protocols used to perform confidential transactions, the physical security to access reserved areas (such as computing server installations).
The information gathered by the questionnaires is used to evaluate whether there are risky practices within the business processes of the company, as well as to propose amendments to these practices.
The ultimate goal of the risk assessment is to determine a degree of compliance to existing standards.
By analyzing the questionnaires, ORCA detects risky practices, proposes solutions and triggers further assessment processes to ensure that the solutions have been implemented.

>>>>

To validate the applicability of our method we have developed a case study concerning risk assessment for financial companies as implemented by the ORCA System\footnote{The ORCA System is a trademark of GCP Global (www.gcpglobal.com).}.
Risk assessment is implemented by an interactive business process based on the exchange of a series of questionnaires intended to evaluate the risks implied by the client business practice.
Examples of business practices are: the conditions and protocols used to perform confidential transactions, the physical security to access reserved areas (such as computing server installations).
The information gathered by the questionnaires is used to evaluate whether there are risky practices within the business processes of the company, as well as to propose amendments to these practices.
The ultimate goal of the risk assessment is to determine a degree of compliance to existing standards.
By analyzing the questionnaires, ORCA detects risky practices, proposes solutions and triggers further assessment processes to ensure that the solutions have been implemented.

### 2015-03-31 JournalToSubmit2015/modelingWithPISODM.tex
first version for IJSEKE

In this section we present \pisodm, an MDD-based methodology for building service compositions with non-functional requirements. 
Our work provides meta-models for modeling functional and non-functional requirements organized in three levels (Figure~\ref{fig:piSOD-M}): CIM (\textit{Computational Independent Models}), PIM (\textit{Plat\-form Independent Models}) and PSM (\textit{Platform Specific Models}). 
Given  high-level models specified at the CIM level, \pisodm proposes the semi-automatic refinement of these models, for the generation of a set of models at the other levels of abstraction.
The refinement process is driven by transformation rules specified between the meta-models.

@@

The e$^3$value  model is used to model the (economic) value exchange among the actors involved in an application. 
However, the e$^3$val\-ue model does not help to understand the business process of the application nor the conditions in which the different steps of this process are executed. 
In order to face this shortcoming, our methodology proposes the BPMN meta-model as a tool for modelling this aspect of the application.

@@

The CIM level models are the basis for the development of the application. 
The information presented at this level is (manually) refined into PIM-level models of the methodology \pisodm described next.
Notice that, at this level, the models are (informally) used to describe both functional and non-functional requirements.

@@

This level focuses on the system functionality, hiding the details of any particular platform.
The specification defines those parts of the system that do not change from one platform to another. 
Our methodology defines three PIM-level meta-models: \textit{$\pi$-UseCase}, \textit{$\pi$-Ser\-vice\-Pro\-cess} and \textit{$\pi$-ServiceComposition}.

>>>>

In this section we present \pisodm, an MDD-based method for building service compositions with non-functional requirements.
Our work provides meta-models for modeling functional and non-functional requirements organized in three levels (Figure~\ref{fig:piSOD-M}): CIM (\textit{Computational Independent Models}), PIM (\textit{Plat\-form Independent Models}) and PSM (\textit{Platform Specific Models}).
Given  high-level models specified at the CIM level, \pisodm proposes the semi-automatic refinement of these models, for the generation of a set of models at the other levels of abstraction.
The refinement process is driven by transformation rules specified between the meta-models.

@@

The e$^3$value  model is used to model the (economic) value exchange among the actors involved in an application.
However, the e$^3$val\-ue model does not help to understand the business process of the application nor the conditions in which the different steps of this process are executed.
In order to face this shortcoming, our method proposes the BPMN meta-model as a tool for modelling this aspect of the application.

@@

The CIM level models are the basis for the development of the application.
The information presented at this level is (manually) refined into PIM-level models of the method \pisodm described next.
Notice that, at this level, the models are (informally) used to describe both functional and non-functional requirements.

@@

This level focuses on the system functionality, hiding the details of any particular platform.
The specification defines those parts of the system that do not change from one platform to another.
Our method defines three PIM-level meta-models: \textit{$\pi$-UseCase}, \textit{$\pi$-Ser\-vice\-Pro\-cess} and \textit{$\pi$-ServiceComposition}.

### 2015-03-31 JournalToSubmit2015/relworks.tex
first version for IJSEKE

@@

\bigskip
Most works agree on distinguishing three points of view, namely the point of view of the organization (\textit{Business} view), of the individual service providers (\textit{Service} view) and of the composition designer (\textit{System} view).
The Business view is concerned with the business logic (\textit{i.e.}, an abstract level of tasks, defined by the guidelines and constraints imposed by the organization).
Service and System views guides the implementation of the software solution: The Service level is related to the building blocks of the application.
It may use web services provided by third party sources.
The System level deals with the coordination of services, to implement the business logic.
Our proposal, shown in the next section, follows these ideas.

>>>>

Beside the existence of these works, it is important to remark that authors continue emphasizing the limited attention of the model-driven proposals for service development to non-functional requirements~\cite{Ameller201542,Ameller2010}. As Ameller et al. realize in a recent published literature review \textit{"Not many papers that presented a generic approach to deal with NFRs"}. Authors also remark in such study the lacks of industrial use case~\cite{Ameller201542}. Our method to develop service-oriented applications aims to cover this lack of proposals dealing with non-functional requirements and gives support in addition to real need presented in a company related to the risk assessment industry.

@@

%\bigskip
%*****OJO**** QUITAR!! ESTA DISTINCION NO APARECE EN REALIDAD EN EL PAPER NO???, YO QUITARIA SIN MAS ESTE UTLIMO PARRAFO SALVO QUE LE VEAN SENTIDO****
%Most works agree on distinguishing three points of view, namely the point of view of the organization (\textit{Business} view), of the individual service providers (\textit{Service} view) and of the composition designer (\textit{System} view).
%The Business view is concerned with the business logic (\textit{i.e.}, an abstract level of tasks, defined by the guidelines and constraints imposed by the organization).
%Service and System views guides the implementation of the software solution: The Service level is related to the building blocks of the application.
%It may use web services provided by third party sources.
%The System level deals with the coordination of services, to implement the business logic.
%Our proposal, shown in the next section, follows these ideas.

### 2015-03-31 JournalToSubmit2015/transformationRules.tex
first version for IJSEKE

A {\sf Constraint} associated to a {\sf Use Case}  is transformed into an  {\sf Assertion}.  
The set of resulting assertions  are grouped into a {\sf Contract}.
Constraints are transformed according to their type:
{\sc Business} constraints and {\sc Value} constraints with the {\sf  isExceptionalBehaviour} attribute set to false are transformed into {\sf Assertion}s;
{\sc Value } constraints with the {\sf  isExceptionalBehaviour} attribute set to true are transformed into {\sf Exceptional behavior}s.

In order to transform constraints of type {\sf Value Constraint}, the designer must specify thresholds to be associated to the assertions of a contract.
By default, value constraints are transformed into pre-conditions and business constraints are transformed into post-conditions.

@@

The details of these transformations are not included here (due to space restrictions).
They can be found in~\cite{SouzaNeto:2012}.

\begin{example}[To Publish Music \textit{(cont)}]\label{ex:toPublicMusicT1}
The rules presented above have been applied to the model in Figure~\ref{fig:piUseCaseModel}, in order to obtain the $\pi$-Service Process model for our running example (Figure~\ref{fig:CIM:serviceprocess}).

The ``listen music'' use case is transformed into a Service Action. 
This Action  represents a Spotify service function that can be invoked to play the music. 
For the ``publish music'' use case,  constraints are transformed into a set of assertions that are grouped into a Contract ({\sf ``publishMusicContract''}) associated to the Action {\sf ``publishMusic''}. 
The ``download music'' use case  includes the payment process to buy the music. 
Thus, these use cases  are transformed into {\sf Actions}, and a {\sf Service Activity} that aggregates these {\sf Actions}.   
This \textsf{Service Activity} is transformed into a sequence flow on the $\pi$-service process model.
%(as shown in figure \ref{fig:transformation-example-include}). 
The same rule is applied to the ``publish music'' use case, which has two extended use cases, to publish on Twitter and Facebook.
 \end{example}

@@

{\sf Actions} and {\sf Service activity} of a $\pi$-Service Process model are transformed into their homonym concepts of the $\pi$-Service Composition model.

@@

\begin{example}[To Publish Music \textit{(cont)}]\label{ex:toPublicMusicT5}
Considering the example scenario, the model in Figure~\ref{fig:PiServiceCompositionModel} was obtained by applying the rules above to the model depicted by Figure~\ref{fig:CIM:serviceprocess}.  
The ``securityLoginPolicy'' consists of a set of {\sf Rules} that were transformed from the {\sf Assertions} in $\pi$-service process model. 
The information about Facebook and Spotify (both of them {\sf Business
Collaborators}) come from entity of type {\sc Package} in the $\pi$-use case
model.
\end{example}

@@

This section describes the PIM to PSM transformations from a $\pi$-Service composition model to a $\pi$-PEWS model. 
We distinguish two groups of rules: \textit{(i)} those transforming service composition entities into workflows; and \textit{(ii)} those that transform  A-Policies into Contracts.

Single actions (represented by {\sf Action} and {\sf
Action:name}) are transformed into individual service {\sf Operations}.
Complex actions (represented by {\sf ServiceActivity}  and  {\sf
ServiceActivity:name}) are transformed into (named) composite
operations that defines a (named) workflow of the application.
Composition patterns expressed using the operators {\sc\em merge, decision, fork and join} are transformed into their corresponding workflows of the $\pi$-PEWS model.

@@

The A-Policies defined for the entities of a $\pi$-service composition model are transformed into {\sf A-policy} entities, named according to the names expressed in the source model.
The transformation of the rules expressed in a $\pi$-service composition is guided by the event types associated to these rules. 
The variables associated to an A-Policy expressed in a $\pi$-service composition model as {\sf $<$Variable:name, Variable:type$>$} are transformed into entities  {\sf Variable} with attributes {\sf Name} and {\sf Type} directly specified from the elements {\sf Variable:name} and {\sf Variable:type} of a $\pi$-service composition model.

@@

\subsection{Implementation}
We have developed tools for aiding the user to define and transform the models for all the levels, except for the CIM into PIM level, which should be manually performed.

Our tool is implemented as a series of Eclipse plug-ins: 
\begin{itemize}
\item 	We  used the Eclipse Modeling Framework (EMF)\footnote {The EMF project is a modeling framework and code generation facility for building tools and other applications based on a structured data model.}   for implementing the  $\pi$-Service Composition and $\pi$-{\sc Pews}  meta-models. 
Then, from these meta-models, we  developed plug-ins to support their graphical representation.

@@

\item 	We  used Acceleo\footnote{http://www.acceleo.org/pages/home/en} for implementing  the code generation plug-in for generating executable code. 
It takes a $\pi$-PEWS model and generates the code to be executed by the {\em
A-Policy} based service composition execution environment.
\end{itemize}

>>>>

A {\sf Constraint} associated to a {\sf Use Case}  is transformed into an  {\sf Assertion}.  
The set of resulting assertions  are grouped into a {\sf Contract}.
Constraints are transformed according to their type:
{\sc Business} constraints and {\sc Value} constraints with the {\sf  isExceptionalBehaviour} at\-tri\-bute set to false are transformed into {\sf Assertion}s;
{\sc Value } constraints with the {\sf  isExceptionalBehaviour} attribute set to true are transformed into {\sf Exceptional behavior}s.

In order to transform constraints of type {\sf Value Constraint}, the designer specifies thresholds to be associated to the assertions of a contract.
By default, value constraints are transformed into pre-conditions and business constraints are transformed into post-conditions.

@@

Details of these transformations are not included here (due to space restrictions).
They can be found in~\cite{SouzaNeto:2012}.

\begin{example}[To Publish Music \textit{(cont)}]\label{ex:toPublicMusicT1} \\
The transformation rules have been applied to the model in Figure~\ref{fig:piUseCaseModel}, in order to obtain the $\pi$-Service Process model for our running example (Figure~\ref{fig:CIM:serviceprocess}).

The ``listen music'' use case is transformed into a Service Action. 
This Action  represents a Spotify service function that can be invoked to play the music. 
For the ``publish music'' use case,  constraints are transformed into a set of assertions that are grouped into a Contract ({\sf ``publishMusicContract''}) associated to the Action {\sf ``publishMusic''}. 
The use case ``download music''  includes the payment process to buy the music. 
Thus, these use cases  are transformed into {\sf Actions}, and a {\sf Service Activity} that aggregates these {\sf Actions}.   
This \textsf{Service Activity} is transformed into a sequence flow on the $\pi$-service process model.
%(as shown in figure \ref{fig:transformation-example-include}). 
The same rule is applied to the ``publish music'' use case, which has two extended use cases, to publish on Twitter and Facebook.
 \end{example}

@@

{\sf Actions} and {\sf Service activity} of a $\pi$-Service Process model are transformed into their hom\-onym concepts of the $\pi$-Service Composition model.

@@

\begin{example}[To Publish Music \textit{(cont)}]\label{ex:toPublicMusicT5}\\
Considering the example scenario, the model in Figure~\ref{fig:PiServiceCompositionModel} was obtained by applying the rules above to the model depicted by Figure~\ref{fig:CIM:serviceprocess}.  
The ``securityLoginPolicy'' consists of a set of {\sf Rules} that were transformed from the {\sf Assertions} in $\pi$-service process model. 
The information about Facebook and Spotify (both of them {\sf Business
Collaborators}) come from entity of type {\sc Package} in the $\pi$-use case
model.
\end{example}

@@

This section describes the PIM to PSM transformations from a $\pi$-Service composition mod\-el to a $\pi$-PEWS model. 
We distinguish two groups of rules: \textit{(i)} those transforming service composition entities into workflows; and \textit{(ii)} those transforming  A-Policies into Contracts.

Single actions ({\sf Action} and {\sf
Action:name}) are transformed into individual service {\sf Operations}.
Complex actions (rep\-re\-sented by {\sf ServiceActivity}  and  {\sf
ServiceActivity:name}) are transformed into (named) composite
operations that defines a (named) workflow of the application.
Composition patterns expressed using  {\em merge, decision, fork} and {\em join} are transformed into their corresponding workflows of the $\pi$-PEWS model.

@@

A-Policies defined for the entities of a $\pi$-service composition model generate {\sf A-policy} entities, named according to the names expres\-sed in the source model.
The transformation of the rules expres\-sed in a $\pi$-service composition is guided by the event types associated to these rules. 
The variables associated to an A-Policy expressed in a $\pi$-service composition model as {\sf $<$Variable:name, Variable:type$>$} are transformed into  {\sf Variable}, with  {\sf Name} and {\sf Type} attributes specified from the elements {\sf Variable: name} and {\sf Variable:type} of a $\pi$-service composition model.

@@

\subsection{Implementation}
We have developed tools for aiding the user to define and transform the models for all the levels, except for the CIM into PIM level, which should be manually performed.
Our tool is implemented as a series of Eclipse plug-ins: 
\begin{itemize}
\item 	We  used the Eclipse Modeling Framework (EMF)\footnote {The EMF project is a modeling framework and code generation facility for building tools and other applications based on a structured data model.}   to implement the  $\pi$-Service Composition and $\pi$-{\sc Pews}  meta-models. 
From these meta-models, we  developed plug-ins to support their graphical representation.

@@

\item 	We  used Acceleo\footnote{http://www.acceleo.org/pages/home/en} to implement  the code generation plug-in to generate executable code. 
It takes a $\pi$-PEWS model and generates the code to be executed by the {\em
A-Policy} based service composition execution environment.
\end{itemize}

### 2015-03-31 JournalToSubmit2015/transformations.tex
first version for IJSEKE

\begin{figure}[t]
\centering{
\includegraphics[width=0.96\textwidth]{figs/Mapping-1a.png}
}
\caption{ $\pi$-SCM to $\pi$-{\sc Pews} transformations.}
\label{fig:transformationsA}
\end{figure}

@@

As shown in Figures~\ref{fig:transformationsA} and~\ref{fig:transformationsB}, for an event of type {\sc\em Pre} the corresponding transformed rule is of type {\sc Precondition}; for an event of type {\sc\em Post} the corresponding transformed rule is of type {\sc Postcondition}; finally, for an event of type {\sc\em TimeRestriction} the corresponding transformed rule is of type {\sc Time}. 
The condition inside a $\pi$-SCM rule ({\sc\em Rule:condition}) is transformed into a class {\sc\em Condition:expression} where the attributes of the expression are transformed into elements of type {\sc Attribute}.

>>>>

@@

As shown in Figures~\ref{fig:transformationsB} and~\ref{fig:transformationsC}, for an event of type {\sc\em Pre} the corresponding transformed rule is of type {\sc Precondition}; for an event of type {\sc\em Post} the corresponding transformed rule is of type {\sc Postcondition}; finally, for an event of type {\sc\em TimeRestriction} the corresponding transformed rule is of type {\sc Time}. 
The condition inside a $\pi$-SCM rule ({\sc\em Rule:condition}) is transformed into a class {\sc\em Condition:expression} where the attributes of the expression are transformed into elements of type {\sc Attribute}.
